\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-09-20 notes/2loopNLSM.tex

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c .
\end{equation}
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of YZ-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic 4-point 2-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams {d[p[1], p[3]], d[p[1], p[6]], d[p[1],
% p[7]], d[p[2], p[3]], d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]],
% d[p[3], p[7]], d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-Abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
In other words, \emph{J-theory and semi-Abelian YM are identical}.
By calculating an off-shell 4-point correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
As might be expected, the kinematic algebra is the diffeomorphism algebra.
For details about external states and how to extract NLSM (a scalar theory) from semi-YM (a vector theory) we refer the reader to \cite{Cheung:2021zvb}.
Note that the different external particles break manifest Bose symmetry.

>>>>

We can achieve this construction by splitting the gauge group of the
full Yang-Mills action from $G$ to $G_L \times G_R$, where the full
non-abelian gauge symmetry is restored only when we identify
$G_L = G_R$. However, if we break the left gauge group down to
$G_L \rightarrow U(1)^{N^2}$ and take the right gauge group to
$G_R \rightarrow U(N)$, then we are left with a semi-abelian theory
that is gauge invariant in the preferred $U(1)$-directions. We call
this manifestly color-dual theory \textit{semi-abelian Yang-Mills},
\begin{eBox}
\begin{equation}
\label{eq:semiYM}
\mathcal{L}^{\text{semi-YM}} = -\frac{1}{2}\text{tr}\left[\bar{F}_{\mu\nu}F^{\mu\nu}\right]
\end{equation}
\end{eBox}
where
\begin{equation}
\bar{F}^a_{\mu\nu} = \partial_\mu \bar{A}_\nu^a - \partial_\nu \bar{A}_\mu^a
\end{equation}
\begin{equation}
F^a_{\mu\nu} = \partial_\mu A_\nu^a - \partial_\nu A_\mu^a + f^{abc} A_\mu^b A_\nu^c
\end{equation}
and $A_\mu$ is in Lorenz gauge.
Here we can think of the abelian vector as a background field that
sources on-shell $A_\mu$ currents. Indeed, as we describe in Appendix
\cref{sec:CKLagrangians}, semi-abelian Yang-Mills theory is at the
heart of YZ-theory \cite{Cheung:2016prv}, J-theory
\cite{Cheung:2020djz,Cheung:2021zvb}, and self-dual Yang-Mills
\cite{Monteiro2011pc}.  Specifically, semi-Abelian YM is simply a
clever reinterpretation of the Lagrangian obtained by integrating an
auxiliary field into the J-theory equations of motion.  The Feynman
rule for the cubic vertex is
\begin{equation}
  \cubic{hgrey0}{fermion2}{}{fermion2}{}{fermion2}{}= (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3).
\end{equation}
The kinetic term mixes the two gauge fields and has the standard
Feynman rule of $i/k^2$.  We can use the Ward identity of
\cref{eq:GIrelA} to check that the amplitudes are indeed gauge
invariant when the abelian polarizations are taken to be longitudinal,
$\bar{\varepsilon} \rightarrow k$.

@@

Ignoring tadpoles and BEL graphs, there are 14 cubic 4-point 2-loop
graphs (see \cref{fig:MaxCuts}) which are related to each other via 21
Jacobi relations.  As mentioned in \cref{sec:bootstrap},
color-kinematics duality ensures that the numerator of every graph can
be expressed in terms of a basis of the double box and penta-triangle,
\begin{equation}
\label{eq:jacobi-basis}
\NLSMBasisDoubleBox \text{ and } \NLSMBasisPentaTriangle .
\end{equation}
For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of
these graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
\{ k_{13}, k_{23}, k_{15}, k_{16}, k_{25}, k_{26}, k_{35}, k_{36}, k_{55}, k_{56}, k_{66} \},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to equations with sixteen thousand parameters so a
  different method is desirable.  One possibility is to use a solver
  designed for integral reduction.  However, these solvers typically
  assume that the matrix has a very low density which is not the case
  for equations coming from NLSM which typically have a density of a
  few percent.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions), there is no such
  separation of sectors in a scalar theory where any product of
  momenta can interfere with any other.  The NLSM constraint equations
  were solved using a custom solver, \texttt{FiniteFieldSolve},
  designed to solve large linear systems of arbitrary density.
  \texttt{FiniteFieldSolve} will be released publicly shortly.\ace{I
    want to rephrase this slightly, but need to think about it a bit
    more}} Imposing the boundary Jacobi relations eliminates all but
4,473 of the original free parameters.  Further imposing graph
symmetries reduces the number of free parameters to 1,243.

@@

\paragraph{J-theory} There is an alternative color-dual formulation of NLSM that the theory manifestly corresponds to NLSM at least at tree level \cite{Cheung:2021zvb}.
For related formulations of NLSM that do not treat color-kinematics duality see \cite{Freedman:1980us, Slavnov:1971mz}.
The construction begins with a first order formulation of NLSM in terms of chiral currents $J^\mu$.
First order formulation of NLSM in terms of chiral current $J^\mu$.
Two equations.
Two constraints, the field strength should vanish
\begin{equation}
\label{eq:FieldStrengthJ}
F_{\mu\nu}(J) = \partial_\mu J_\nu - \partial_\nu J_\mu + g[J_\mu , J_\nu]=0
\end{equation}
which then implies that $J^\mu$ is pure gauge $J_\mu = U \partial_\mu U^{-1}$.
The second condition is that the chiral current should be in Lorenz gauge
\begin{equation}
\label{eq:LorenzGaugeJ}
\partial_\mu J^\mu=0
\end{equation}
which then yields the NLSM equation of motion $\partial_\mu (U \partial^\mu U^{-1})=0$.
Taking a linear combination of \cref{eq:FieldStrengthJ} and \cref{eq:LorenzGaugeJ} yields
\begin{equation}
\label{eq:JTheoryEOM}
\Box J^\mu +g f^{abc} J^\nu \partial_\nu J^\mu = 0.
\end{equation}
Integrating in an auxiliary field $\bar{J}^\mu$ as a Lagrange multiplier trivially produces the Lagrangian of semi-abelian YM \cref{eq:semiYM} but with the gauge fields renamed to $J_\mu$ and $\bar{J}_\mu$.
Since both theories are in Lorenz gauge this means that \emph{J-theory and semi-abelian YM are identical}.
By calculating an off-shell 4-point correlation function it is possible to check that this Lagrangian automatically satisfies color-kinematics duality.
As might be expected, the kinematic algebra is the diffeomorphism algebra.
For details about external states and how to extract NLSM (a scalar theory) from semi-YM (a vector theory) we refer the reader to \cite{Cheung:2021zvb}.
Note that the different external particles break manifest Bose symmetry.
